### 2019-06-12 1-Preamble/5-Abstract/abstract.tex
Multi-player -> multi-players

% \newpage
% ----------------------------------------------------------------------------
\begin{center}
	\textbf{{\Large Abstract}}
\end{center}
% \vspace*{1cm}
% \vspace*{-5pt}
%
\begin{small}
% Chapter 1
In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access large-scale Cognitive Radio networks, in unlicensed bands and without central control.
We focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
%  (Chapter~\ref{chapter:1})
% Chapter 4
We propose different models of IoT networks, and we show empirically on both numerical simulations and real-world validation the possible gain of our methods, that use Reinforcement Learning.
%  (Chapter~\ref{chapter:4})
% Chapter 2
The different network access problems are modeled as Multi-Armed Bandits (MAB), but we found that analyzing the realistic models was intractable,
% ,  Chapter~\ref{chapter:2}
because proving the convergence of many end-devices playing a collaborative game without communication no coordination is hard, when end-devices all follow random active pattern.
% Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.
% Chapter 5 and 6
The rest of this manuscript thus studies two restricted models, first multi-player bandits in stationary problems, then non-stationary single-player bandits.
%  (Chapter~\ref{chapter:5})
%  (Chapter~\ref{chapter:6})
% Chapter 3
We also detail another contribution, the SMPyBandits open-source Python library for numerical simulations of MAB problems, which covers all the studied models and more.
%  (Chapter~\ref{chapter:3})
\end{small}

>>>>

% \newpage
% ----------------------------------------------------------------------------
\begin{center}
	\textbf{{\Large Abstract}}
\end{center}
% \vspace*{1cm}
% \vspace*{-5pt}
%
\begin{small}
% Chapter 1
In this PhD thesis, we study wireless networks and reconfigurable end-devices that can access large-scale Cognitive Radio networks, in unlicensed bands and without central control.
We focus on Internet of Things networks (IoT), with the objective of extending the devices' battery life, by equipping them with low-cost but efficient machine learning algorithms, to let them automatically improve the efficiency of their wireless communications.
%  (Chapter~\ref{chapter:1})
% Chapter 4
We propose different models of IoT networks, and we show empirically on both numerical simulations and real-world validation the possible gain of our methods, that use Reinforcement Learning.
%  (Chapter~\ref{chapter:4})
% Chapter 2
The different network access problems are modeled as Multi-Armed Bandits (MAB), but we found that analyzing the realistic models was intractable,
% ,  Chapter~\ref{chapter:2}
because proving the convergence of many end-devices playing a collaborative game without communication no coordination is hard, when end-devices all follow random active pattern.
% Indeed, even the number of active devices at any time is highly unpredictable, making the environment not stationary and quickly evolving.
% Chapter 5 and 6
The rest of this manuscript thus studies two restricted models, first multi-players bandits in stationary problems, then non-stationary single-player bandits.
%  (Chapter~\ref{chapter:5})
%  (Chapter~\ref{chapter:6})
% Chapter 3
We also detail another contribution, the SMPyBandits open-source Python library for numerical simulations of MAB problems, which covers all the studied models and more.
%  (Chapter~\ref{chapter:3})
\end{small}

### 2019-06-12 2-Chapters/1-Chapter/chapter1.tex
Multi-player -> multi-players

@@

    \item
    We formalize the multi-player bandit model, and we introduced three variants, in Chapter~\ref{chapter:5}.
    For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
    as well as extensive numerical experiments to demonstrate its good performance in comparison with the rest of the literature.
    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-player bandits, as some recent research works built up on our results.

    \item
    We also give a detailed literature review of the different extensions of the multi-player MAB model, which we believe was never written before.
    % were  (ALT 2018, and 4-8 works inspired by our article since then). State-of-the-art with our algorithm MCTopM + klUCB for multi-player bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"

@@

% ----------------------------------------------------------------------------
\section{List of publications and other experiences}
\label{sec:1:listPublications}

% =============================================================================
\subsection{Publications in international conferences with proceedings}

@@

% =============================================================================
\subsection{Demonstration in international conferences}

@@

% =============================================================================
\subsection{French language conference with proceedings}
% \subsection{Submitted works}

@@

% =============================================================================
\subsection{In progress works waiting for a new submission}

@@

% =============================================================================
\subsection{Other works}

\begin{itemize}
\item
    \emph{A Note on the Ei Function and a Useful Sum-Inequality},
    by \textbf{Lilian Besson},
    February $2018$,
    \href{https://HAL.Inria.fr/hal-01847480}{\texttt{HAL.Inria.fr/hal-01847480}}.

% =============================================================================
\subsection{Presentations in seminars and conferences}

\textbf{Seminars.}
    I gave some presentations in the following events:
    SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
    SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
    as well as
    for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
    for the brown-bag seminar at ENSAI in Bruz in January $2018$,
    for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
    for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in June $2019$.

\textbf{Tutorial.}
    I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
    with Pierre Haessig (see \href{https://pierreh.eu/}{\texttt{pierreh.eu}}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.

\textbf{Training.}
    I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
    We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
    and I gave about $12$ training sessions, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.

% =============================================================================
\subsection{Other experiences}

\textbf{Conferences.}
    I attended the following conferences:
	\emph{International Conference on Communication} ICC (Paris), May $2017$,
    \emph{Conference on Learning Theory} COLT (Amsterdam), July $2017$,
    \emph{Conference on Cognitive Radio Oriented Wireless Networks} CROWNCOM (Lisboa), September $2017$,
    \emph{Conference on Algorithmic Learning Theory} ALT (Lanzarote), April $2018$,
    \emph{Wireless Communication and Networking Conference} WCNC (Barcelona), April $2018$,
	\emph{International Conference on Telecommunication} ICT (Saint-Malo), June $2018$,
    \emph{Wireless Communication and Networking Conference} WCNC (Marrakech), April $2019$,
    \emph{Colloque francophone de traitement du signal et des images} GRETSI (Lille), August $2019$.

\textbf{Seminars.}
    I attended the following seminars:
	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$.

\textbf{Responsabilities.}
    I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
    I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
    I also helped for the organization and presented another poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.

\textbf{Reviews.}
	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

\textbf{System Administrator.}
	maintaining our workstations running GNU/Linux and Windows
	at SCEE team,
	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
	from January $2017$ to September $2018$.

@@

\TODOL{Open source the repository as soon as I defended my thesis!}

>>>>

\paragraph{Applications for Opportunistic Spectrum Access.}
% Detail the previous work from our team SCEE on bandits + OSA : Wassim, Navik

The focus of this work is on cognitive radio and IoT networks, where arms can represent wireless orthogonal channels, but more generally any resource characterizing the communication between a wireless device and a gateway (\eg, spreading factor for LoRa \cite{KerkoucheAlami18}, power allocation for NOMA etc). In cognitive radio using centralized supervision, for instance if the gateway can decide the allocation of devices to resources, MAB can also be used to let the gateway explore different allocations and learn by itself a good allocation, see for instance this article that consider 5G-like networks with small cells \cite{Maghsudi16}.

Previous works of our SCEE team showed that MAB can be used to model the problem of spectrum access for a secondary user accessing a licensed spectrum.
In this model, arms represent a finite set of orthogonal channels, \ie, different frequency bands in a licensed spectrum.
In the model with sensing, the samples $Y_{k,t}$ represents the feedback obtained by the CR-equipped device after sensing the channel $k$ at time $t$.
A reward of $r(t) = 1$ indicates that no Primary User was sensed, while a reward of $r(t)=0$ indicates that the channel $k$ is busy at time $t$ and no uplink message can be sent.
%
This model was first introduced by Wassim Jouini during his PhD thesis,
in \cite{Jouini09} and later studied in both \cite{Jouini10,Jouini12}.
Proof-of-concepts using real-world radio hardware was first propsed in \cite{MoyWSR2014,RobertSDR2014}.

\TODOL{Je sais pas à quel point je veux parler de la thèse de Navik? Peut-être que j'en aurai dit assez au Chapitre 1, et que là ça suffira..}
In a second PhD thesis supervised by Christophe Moy, Navikkumar Modi studied the impact
of using MAB algorithms to optimize channel selection
on the battery life of a wireless device.
On the one hand, running a MAB algorithm such as UCB-like algorithms was proved to be useful and can bring significant improvement in terms of successful transmission rates, directly increasing the battery life of the device.
On the other hand, classical MAB algorithms tend to switch arms a lot of times, especially in the beginning of the learning process, and this induces a lot of dynamical hardware reconfiguration for the wireless device, as selecting a different channel requires a change in the radio hardware used by the device.
Each hardware reconfiguration costs energy for the device, and quickly switching algorithms will lead to a reduction of the battery life.
The tradeoff between the two aspects is studied empirically in
\cite{modiDemo2016}.
%
Another interesting works is \cite{Modi17QoS}.

\TODOL{FIXME Finir ce blabla rapidement ? Ou alors ce sera déjà fait en intro/chapitre1 ??}

@@

    \item
    We formalize the multi-players bandit model, and we introduced three variants, in Chapter~\ref{chapter:5}.
    For the case with sensing information, we propose two new algorithms, and we give an analysis for our algorithm \MCTopM{} to show it is order-optimal,
    as well as extensive numerical experiments to demonstrate its good performance in comparison with the rest of the literature.
    Our work \cite{Besson2018ALT} also gave a new impulse on research on multi-players bandits, as some recent research works built up on our results.

    \item
    We also give a detailed literature review of the different extensions of the multi-players MAB model, which we believe was never written before.
    % were  (ALT 2018, and 4-8 works inspired by our article since then). State-of-the-art with our algorithm MCTopM + klUCB for multi-players bandits "with sensing", empirical state-of-the-art with our simple (but wrong) "selfish" approach in case of "no sensing"

@@

% FIXME
\newpage

% ----------------------------------------------------------------------------
\section{List of publications}
\label{sec:1:listPublications}

% =============================================================================
\paragraph{Publications in international conferences with proceedings}

@@

% =============================================================================
\textbf{Demonstration in international conferences}

@@

% =============================================================================
\textbf{French language conference with proceedings}
% \textbf{Submitted works}

@@

% =============================================================================
\textbf{In progress works waiting for a new submission}

@@

% % =============================================================================
% \textbf{Other works}

% \begin{itemize}
% \item
%     \emph{A Note on the Ei Function and a Useful Sum-Inequality},
%     by \textbf{Lilian Besson},
%     February $2018$,
%     \href{https://HAL.Inria.fr/hal-01847480}{\texttt{HAL.Inria.fr/hal-01847480}}.

% \end{itemize}

% % =============================================================================
% \textbf{Presentations in seminars and conferences}

% \textbf{Seminars.}
%     I gave some presentations in the following events:
%     SequeL team seminar at Inria Lille in September and December $2017$, and June $2019$;
%     SCEE team seminar at CentraleSupélec, Rennes campus, in October $2017$, February $2018$ and June $2019$;
%     as well as
%     for the GDR ISIS day held in Issy-les-Moulineaux on November $2017$,
%     for the brown-bag seminar at ENSAI in Bruz in January $2018$,
%     for the weekly seminar at CMAP lab at École Polytechnique in October $2018$,
%     for the weekly seminar of the PANAMA project-team at IRISA / Inria Rennes in June $2019$.

% \textbf{Tutorial.}
%     I gave a tutorial on the Julia language, at IETR seminar in Vannes in June $2018$,
%     with Pierre Haessig (see \href{https://pierreh.eu/}{\texttt{pierreh.eu}}), see \href{https://HAL.Inria.fr/cel-01830248}{\texttt{HAL.Inria.fr/cel-01830248}}.

% \textbf{Training.}
%     I was also in charge of ``GouTP'' training sessions for about $30$ PhD students at CentraleSupélec Rennes.
%     We had a lot of various $1$h training sessions between January $2017$ and June $2019$,
%     and I gave about $12$ training sessions, on various topics including Python, \texttt{git}, HAL and arXiv, the Julia language and Bib\TeX{}.

% % =============================================================================
% \textbf{Other experiences}

% \textbf{Conferences.}
%     I attended the following conferences:
% 	\emph{International Conference on Communication} ICC (Paris), May $2017$,
%     \emph{Conference on Learning Theory} COLT (Amsterdam), July $2017$,
%     \emph{Conference on Cognitive Radio Oriented Wireless Networks} CROWNCOM (Lisboa), September $2017$,
%     \emph{Conference on Algorithmic Learning Theory} ALT (Lanzarote), April $2018$,
%     \emph{Wireless Communication and Networking Conference} WCNC (Barcelona), April $2018$,
% 	\emph{International Conference on Telecommunication} ICT (Saint-Malo), June $2018$,
%     \emph{Wireless Communication and Networking Conference} WCNC (Marrakech), April $2019$,
%     \emph{Colloque francophone de traitement du signal et des images} GRETSI (Lille), August $2019$.

% \textbf{Seminars.}
%     I attended the following seminars:
% 	\emph{Workshop Learn with Earning} (Rotterdam), May $2018$,
% 	\emph{Workshop on Optimization and Learning} (Toulouse), September $2018$,
% 	\emph{European Workshop on Reinforcement Learning} (Lille), October $2018$.

% \textbf{Responsibilities.}
%     I was the president of the PhD Students Association of IETR lab (ADDI, \texttt{addi.asso.insa-rennes.fr}) in $2017$.
%     I was notably in charge of organizing the PhD Students day at Rennes, in June $2017$, with about $350$ people, and presentation of a research poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013839}{\texttt{HAL.Inria.fr/hal-02013839}}}
%     I also helped for the organization and presented another poster\footnote{~See \href{https://HAL.Inria.fr/hal-02013847}{\texttt{HAL.Inria.fr/hal-02013847}}} at the ``IETR : Interagir Évaluer Transmettre Réunir'' seminar, held in Vannes in June $2018$.

% \textbf{Reviews.}
% 	for \emph{European Workshop on Reinforcement Learning}\footnote{~EWRL 2018, see \href{https://ewrl.wordpress.com/ewrl14-2018}{\texttt{ewrl.wordpress.com/ewrl14-2018}}} (Lille), in October $2018$, I reviewed $5$ papers.
% 	I helped colleagues for articles submitted at international conferences \emph{AISTATS} $2019$, \emph{NeurIPS} $2017$ et $2018$, and \emph{COLT} $2018$ and $2019$, for about 10 \emph{reviews}.

% \textbf{System Administrator.}
% 	maintaining our workstations running GNU/Linux and Windows
% 	at SCEE team,
% 	and our cognitive radio test-bed using USRP cards and the GNU Radio software,
% 	from January $2017$ to September $2018$.

@@

% FIXME
% \TODOL{Open source the repository as soon as I defended my thesis!}
